---
title: "dplyr"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## library 

```{r}
library(dplyr)
```
## building datasets
```{r}
mydata = read.csv("income.csv")

```

## Selecting Random N Rows
```{r}
sample_n(mydata,3)
```

## Selecting Random Fraction of Rows
```{r}
sample_frac(mydata,0.1)
```


## Remove Duplicate Rows based on all the variables (Complete Row), one variable or multiple variables
```{r}
x1 = distinct(mydata)
dim(mydata)
dim(x1)

x2 = distinct(mydata, Index, .keep_all= TRUE)
x2 = distinct(mydata, Index, Y2010, .keep_all= TRUE)
```

## Select function
```{r}
# select variables
mydata2 = select(mydata, Index, State:Y2006)
mydata2
```

```{r}
# drop variables
mydata_d = select(mydata, -c(Index,State))
mydata_d
```

```{r}
# select or drop with certain features
mydata3 = select(mydata, starts_with("Y"))
mydata3 = select(mydata, -starts_with("Y"))
```

Helpers	Description
starts_with()	Starts with a prefix
ends_with()	Ends with a prefix
contains()	Contains a literal string
matches()	Matches a regular expression
num_range()	Numerical range like x01, x02, x03.
one_of()	Variables in character vector.
everything()	All variables.

```{r}
mydata4 = select(mydata, contains("I"))
mydata5 = select(mydata, State, everything())
```


## rename( ) Function
```{r}
# we are renaming 'Index' variable to 'Index1'
mydata6 = rename(mydata, Index1=Index)
```

## dplyr

```{r cars}
summary(cars)
```


