---
title: "survival_assumption"
author: "Jerome"
date: "5/27/2020"
output: html_document
---

```{r}
library("survival")
library("survminer")

res.cox <- coxph(Surv(time, status) ~ age + sex + wt.loss, data =  lung)
res.cox
```

In principle, the Schoenfeld residuals are independent of time. A plot that shows a non-random pattern against time is evidence of violation of the PH assumption.

The proportional hazard assumption is supported by a non-significant relationship between residuals and time, and refuted by a significant relationship.

```{r}
test.ph <- cox.zph(res.cox)
test.ph
```

```{r}
?cox.zph
```

```{r}
ggcoxzph(test.ph)

# The pattern looks fairly symmetric around 0.
```

```{r}
ggcoxzph(test.ph, var = "age")

```


```{r}
ggcoxzph(test.ph, var = "sex")

```

# Testing influential observations

Specifying the argument type = “dfbeta”, plots the estimated changes in the regression coefficients upon deleting each observation in turn; likewise, type=“dfbetas” produces the estimated changes in the coefficients divided by their standard errors.

```{r}
ggcoxdiagnostics(res.cox, type = "dfbeta",
                 linear.predictions = FALSE, ggtheme = theme_bw())
```




```{r}
ggcoxdiagnostics(res.cox, type = "deviance",
                 linear.predictions = FALSE, ggtheme = theme_bw())
```

