---
title: "c_interface"
author: "Jerome"
date: "5/5/2020"
output: html_document
---

```{r}
#install.packages("inline")
#install.packages("pryr")
library(inline)
library(pryr)
```


http://adv-r.had.co.nz/C-interface.html#calling-c

# short example
```{r}
add <- cfunction(c(a = "integer", b = "integer"), "
  SEXP result = PROTECT(allocVector(REALSXP, 1));
  REAL(result)[0] = asReal(a) + asReal(b);
  UNPROTECT(1);

  return result;
")

add(1, 5)
#> [1] 6
```

# data structures 
```{r}
library(pryr)
#> Registered S3 method overwritten by 'pryr':
#>   method      from
#>   print.bytes Rcpp

sexp_type(10L)
#> [1] "INTSXP"
sexp_type("a")
#> [1] "STRSXP"
sexp_type(T)
#> [1] "LGLSXP"
sexp_type(list(a = 1))
#> [1] "VECSXP"
sexp_type(pairlist(a = 1))
#> [1] "LISTSXP"
```

# PROTECT key words 

At the heart of every C function are conversions between R data structures and C data structures. Inputs and output will always be R data structures (SEXPs) and you will need to convert them to C data structures in order to do any work. This section focusses on vectors because they’re the type of object you’re most likely to work with.

An additional complication is the garbage collector: if you don’t PROTECT an R object you create, the garbage collector will think they are unused and delete them.

The simplest way to create a new R-level object is to use allocVector(). It takes two arguments, the type of SEXP (or SEXPTYPE) to create, and the length of the vector. The following code creates a three element list containing a logical vector, a numeric vector, and an integer vector, all of length four:


```{r}
dummy <- cfunction(body = '
  SEXP dbls = PROTECT(allocVector(REALSXP, 4));
  SEXP lgls = PROTECT(allocVector(LGLSXP, 4));
  SEXP ints = PROTECT(allocVector(INTSXP, 4));

  SEXP vec = PROTECT(allocVector(VECSXP, 3));
  SET_VECTOR_ELT(vec, 0, dbls);
  SET_VECTOR_ELT(vec, 1, lgls);
  SET_VECTOR_ELT(vec, 2, ints);

  UNPROTECT(4);
  return vec;
')
dummy()
#> [[1]]
#> [1] 4.64e-310 4.64e-310 4.64e-310 4.64e-310
#> 
#> [[2]]
#> [1]   NA TRUE TRUE TRUE
#> 
#> [[3]]
#> [1] NA  1  1  1
```

You might wonder what all the PROTECT() calls do. They tell R that the object is in use and shouldn’t be deleted if the garbage collector is activated. (We don’t need to protect objects that R already knows we’re using, like function arguments.)

You also need to make sure that every protected object is unprotected. UNPROTECT() takes a single integer argument, n, and unprotects the last n objects that were protected. The number of protects and unprotects must match. If not, R will warn about a “stack imbalance in .Call”. Other specialised forms of protection are needed in some circumstances:

UNPROTECT_PTR() unprotects the object pointed to by the SEXPs.

PROTECT_WITH_INDEX() saves an index of the protection location that can be used to replace the protected value using REPROTECT().


# memset() funtion   

you may want to loop through each element in the vector and set it to a constant. The most efficient way to do that is to use memset()
```{r}
zeroes <- cfunction(c(n_ = "integer"), '
  int n = asInteger(n_);

  SEXP out = PROTECT(allocVector(INTSXP, n));
  memset(INTEGER(out), 0, n * sizeof(int));
  UNPROTECT(1);

  return out;
')
zeroes(10);
#>  [1] 0 0 0 0 0 0 0 0 0 0
```

# missing values 
```{r}
is_na <- cfunction(c(x = "ANY"), '
  int n = length(x);

  SEXP out = PROTECT(allocVector(LGLSXP, n));

  for (int i = 0; i < n; i++) {
    switch(TYPEOF(x)) {
      case LGLSXP:
        LOGICAL(out)[i] = (LOGICAL(x)[i] == NA_LOGICAL);
        break;
      case INTSXP:
        LOGICAL(out)[i] = (INTEGER(x)[i] == NA_INTEGER);
        break;
      case REALSXP:
        LOGICAL(out)[i] = ISNA(REAL(x)[i]);
        break;
      case STRSXP:
        LOGICAL(out)[i] = (STRING_ELT(x, i) == NA_STRING);
        break;
      default:
        LOGICAL(out)[i] = NA_LOGICAL;
    }
  }
  UNPROTECT(1);

  return out;
')
is_na(c(NA, 1L))
#> [1]  TRUE FALSE
is_na(c(NA, 1))
#> [1]  TRUE FALSE
is_na(c(NA, "a"))
#> [1]  TRUE FALSE
is_na(c(NA, TRUE))
#> [1]  TRUE FALSE

```


# accessing the data 
```{r}

```

